---
title: "ggplot proj"
author: "Yoon Sung Hong"
date: "9/26/2017"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(readr)
library(ggplot2)
library(dplyr)
setwd("~/stat133/stat133-hws-fall17/hw02")
```

## Import the data in R

```{r}
dat <- read.csv("./data/nba2017-player-statistics.csv", header = TRUE,
                colClasses = c("Player" = "character",
                               "Team" = "character",
                               "Experience" = "character",
                                "Position" = "factor",
                                "Salary" = "double"
                                ))
str(dat)
```

```{r}
dat_2 <- read_csv("./data/nba2017-player-statistics.csv",
                  col_types = list(.default = col_integer(),
                                   Player = col_character(), 
                                   Team = col_character(), 
                                   Experience = col_character(),
                                   Position = col_factor(c("C","PF","PG","SF","SG")),
                                   Salary = col_double()))
str(dat_2)
```


>The rest of the 19 columns are already classified as **integer**, as shown from the `str()` function.
>

## Right after importing the data
```{r}
#Replacing R's with 0's and converting the variable to integer
dat$Experience <- as.integer(replace(dat$Experience, grep("R", dat$Experience), "0"))
is.integer(dat$Experience)
```

## Performance of players

```{r}
#Adding variables
MFG <- (dat$FGA) - (dat$FGM)
MFT <- (dat$FTA)-(dat$FTM)
PTS <- (3*(dat$Points3))+(2*(dat$Points2))+(dat$FTM)
REB <- (dat$DREB)+(dat$OREB)
MPG <- (dat$MIN)/(dat$GP)
dat$MFG <- MFG
dat$MFT <- MFT
dat$PTS <- PTS
dat$REB <- REB
dat$MPG <- MPG
#Forming [EFF value](https://en.wikipedia.org/wiki/Efficiency_(basketball))
dat$EFF <- ((dat$PTS)+(dat$REB)+(dat$AST)+(dat$STL)+(dat$BLK)-(dat$MFG)-(dat$MFT)-(dat$TO))/(dat$GP)
#Summary
summary(dat$EFF)
```

```{r}
#Histogram
hist(dat$EFF, xlab = "EFF", xlim = c(-10, 40), main = "Histogram of Efficiency (EFF)", col = "light grey")
```

```{r}
#Top 10 EFF-valued players
dat[order(-dat$EFF),][(1:10), c("Player", "Team", "Salary", "EFF")]
```
```{r}
#Names of players with negative EFF values
dat[dat$EFF < 0, "Player"]
```
```{r}
#Finding correlations
cor(dat$EFF, dat$PTS)
cor(dat$EFF, dat$REB)
cor(dat$EFF, dat$AST)
cor(dat$EFF, dat$STL)
cor(dat$EFF, dat$BLK)
cor(dat$EFF, -dat$MFG)
cor(dat$EFF, -dat$MFT)
cor(dat$EFF, -dat$TO)
```
```{r}
#Displaying correlations in descending order
cor_PTS <-cor(dat$EFF, dat$PTS)
cor_REB <- cor(dat$EFF, dat$REB)
cor_AST <- cor(dat$EFF, dat$AST)
cor_STL <- cor(dat$EFF, dat$STL)
cor_BLK <- cor(dat$EFF, dat$BLK)
cor_MFG <- cor(dat$EFF, -dat$MFG)
cor_MFT <- cor(dat$EFF, -dat$MFT)
cor_TO <- cor(dat$EFF, -dat$TO)
correlations <- data.frame(player.stats = 
                             c("PTS", "REB", "AST", "STL", "BLK", "MFG", "MFT", "TO"),
                           cor = c(cor_PTS, cor_REB, cor_AST, cor_STL, cor_BLK, cor_MFG, cor_MFT, cor_TO))
correlations <- correlations[order(-correlations$cor),]
correlations
#Barplot
barplot(correlations$cor, ylim = c(-1, 1),
        main = "Correlations between Player Stats and EFF", 
        col = c("orange", "grey")[(correlations$cor>0)+1],
        names.arg = correlations$player.stats)
abline(h=0)
```

## Efficiency and Salary

```{r}
#Plotting scatter with lowess using ggplot
ggplot(data = dat, aes(x = EFF, y = Salary)) + geom_point(aes(color = "orange")) + geom_smooth()
```

```{r}
#Plotting scatter with lowess using the more traditional method
plot(dat$EFF, dat$Salary, xlab = "Efficiency", ylab = "Salary")
lines(lowess(dat$EFF, dat$Salary, f = 0.7), col = "Grey", lwd = 3)
```

```{r}
#Correlation coefficient between EFF and salary
cor(dat$EFF, dat$Salary)
```
>Positive correlation between EFF and Salary indicate that generally, higher efficiency from a player is responded through higher salary. The degree to which the two variables correlate is pretty significant, represented by the value of ~0.656.
```{r}
#more "solid" players
player2 <- dat[dat$MPG >= 20,]
#scatterplot
ggplot(data = player2, aes(x = EFF, y = Salary)) + geom_point(aes(color = "orange")) + geom_smooth()

#correlation coefficient
cor(player2$EFF, player2$Salary)
```
>Positive correlation between EFF and Salary (for players with playtime per game over 20 minutes) indicate that generally, higher efficiency from a player is responded through higher salary. The degree to which the two variables correlate is pretty significant, but not as significant as the correlation representation based on all NBA players. The correlation value's significance is represented by the value of ~0.537.  
>
  
    